PRISM protocol v0.3

### 2023-07-21 prism-backend/deprecated/research/article.tex
build: try new build system

In this section we will informally describe a protocol to create and manage DIDs (\textbf{D}ecentralized textbf{Id}entifiers), that also allows to manage the creation, revocation and presentation of verifiable credentials.

@@

In the following steps, we assume the issuer and subjects have already established a secure communication channel.
In order to issue a batch of an arbitrary number $N$ of credentials, the issuer will:
\begin{itemize}
\item Ask to each subject the DID they would like to use for the credential they will receive.
\item The issuer creates the credentials for each subject. It uses his DID and the key id of an \emph{issuing key} to populate
      the credential corresponding fields. On each credential, it adds the corresponding subject DID as one of the claims.
\item The issuer encodes each individual credential using a base64URL encoding.
\item The issuer signs each individual encoded credential with an \emph{issuing key} associated to his publicly known DID. 
      The signing key corresponds to the key id in the credentials' claims.
\item The issuer encodes each signature, and concatenate them to their corresponding credentials using a dot (".") as separator.
      This produces strings of the form $\langle{}encoded\_credentials\rangle{}.\langle{}encoded\_signature\rangle{}$. We call these strings as \emph{signed credentials}.
\item Now, the issuer takes all the signed credentials, and computes a merkle root from them.
\item The issuer creates an \emph{IssueBatch} event that contains the merkle root, and signs the event with an \emph{issuing key} associated to his publicly known DID. 
\textbf{Note}: Today we use the same issuing key for the event signature and the individual credentials signature, but those could be different keys.
\item The issuer attaches the signed event to the metadata of a transaction and sends it to the blockchain.
\item The issuer gives to each user their corresponding signed credential along with their associated merkle inclusion proof.
\end{itemize}

Likewise previous protocol events, all parties will process the transaction once confirmed, validate the event signature and timestamp the merkle root with the time associated to the carrier transaction.

@@

In order to revoke a credentials, issuers have two alternatives:
\begin{enumerate}
\item They revoke all the credentials in a batch.
      This is done by signing a \emph{RevokeBath} event with a \emph{revocation} key associated to the issuer DID.
      The signed event contains the merkle root to revoke.
\item They revoke specific credentials associated to a batch.
      This is done by signing a \emph{RevokeCredentials} event with a \emph{revocation} key associated to the issuer DID. 
      The signed event contains both the merkle root associated to the credentials to revoke, and the hashes of the specific credentials to revoke.
\end{enumerate}

@@

Once they receive credentials from issuers, subjects will present them to interested parties, called \emph{verifiers}. For example, a student may receive a verifiable credential from a university, and would like to present his credential to a potential employer. In our setting, the verifier will be a party following our protocol events from the blockchain. The steps to present and verify a credential are the following:
\begin{itemize}
\item We assume a safe communication channel between the subject and verifier.
      We also assume that the channel can be identified by an identifier $ch$.
\item The subject shares his credential and merkle inclusion proof to the verifier.
\item The verifier then: 
	\begin{itemize}
	\item computes the merkle root from the inclusion proof and the credential hash;
	\item extracts the issuer DID and signing key id from the credential claims,
	\item retrieves from the state he computed from the blockchain, the timestamps associated to the merkle root and the issuing key, 
	\item checks if the merkle root or credential hash has been revoked
	\item validates the issuer signature on the credential, and determines if it was signed at a time when the issuer key was valid.
	\end{itemize}
  \textbf{Note:} The signature of the IssueBatch event has already been verified by the protocol participants at the time of batch publication. 
\item At this point, the verifier knows that the credential was properly signed. Now, he shares a nonce to the subject and asks him to sign it with a key associated to his DID.
\item The subject signs the hash of $nonce || ch$ and returns the signature to the verifier.
\item The verifier now checks the signature and concludes that the credential subject is indeed the person presenting it.
\end{itemize}

@@

In a nutshell, the proposed protocol (see \figref{fig:e2econnector}) is composed
of two phases. In the first phase, the issuer shares a QR code with the holder.
This QR code contains the issuer's public key (or, maybe, its DID), a random
value ($code$, in \figref{fig:e2econnector}), and a session identifier ($token$,
in \figref{fig:e2econnector}). This exchange takes place through a secure (i.e.,
confidential and authenticated) channel, such as physical exchange. Probably
immediately, although also possibly at a later point in time, the prospective
holder sends back her own public key (or, alternatively, her DID) authenticated
with a MAC, where we use the random value
in the QR code as secret key. The issuer uses the session identifier to locate
the corresponding session and, now that it received in an authenticated manner
the public key of the holder, can leverage hybrid encryption techniques to
establish an end-to-end encrypted session, e.g., using ECIES \cite{abr01}.
Ensuring end-to-end encryption and authentication is important at this point, as
all the information exchanged between issuer and holder is routed through the
connector: an entity or set of entities who ensure that asynchronous
communication can take place between issuer and holder. Thus, ECIES is crucial
to ensure the secrecy of the information, end-to-end. Since the holder needs to
have certainty that the data it receives comes from the issuer (and ECIES does
not authenticate the sender), we have the issuer digitally sign the data it
sends through the ECIES-secured channel. Note that we could involve the random
value shared via the QR code for this. But that would force us to re-execute the
first phase of the protocol even when the QR code is compromised after having
completed the first phase of the protocol. Digital signatures using the issuer's
keypair prevent this.

@@

While the protocol is simple, and is based on common cryptographic constructions,
it is easy to get such protocols wrong, as pointed out in numerous occasions
\cite{davi01, kraw01}. Hence, we have used Tamarin%
\footnote{https://tamarin-prover.github.io/ (Last access, October 5th, 2021.)}
to model our protocol\footnote{Source code for the model available here
  \url{https://github.com/input-output-hk/atala-prism/blob/research-tamarin-connector/prism-backend/research/formalizations/connector.spthy}
  \todo{Update to final URL.}}, which is schematized in \figref{fig:e2econnector}

>>>>

In this section we will informally describe a protocol to create and manage DIDs (\textbf{D}ecentralized \textbf{Id}entifiers), that also allows to manage the creation, revocation and presentation of verifiable credentials.

@@

In the following steps, we assume the issuer and subjects have already established a secure communication channel.
In order to issue a batch of an arbitrary number $N$ of credentials, the issuer will:
\begin{itemize}
\item Ask to each subject the DID they would like to use for the credential they will receive.
\item The issuer creates the credentials for each subject. It uses his DID and the key id of an \emph{issuing key} to populate
      the credential corresponding fields. On each credential, it adds the corresponding subject DID as one of the claims.
\item The issuer encodes each individual credential using a base64URL encoding.
\item The issuer signs each individual encoded credential with an \emph{issuing key} associated to his publicly known DID. 
      The signing key corresponds to the key id in the credentials' claims.
\item The issuer encodes each signature, and concatenate them to their corresponding credentials using a dot (".") as separator.
      This produces strings of the form $\langle{}encoded\_credentials\rangle{}.\langle{}encoded\_signature\rangle{}$. We call these strings as \emph{signed credentials}.
\item Now, the issuer takes all the signed credentials, and computes a Merkle root from them.
\item The issuer creates an \emph{IssueBatch} event that contains the Merkle root, and signs the event with an \emph{issuing key} associated to his publicly known DID. 
\textbf{Note}: Today we use the same issuing key for the event signature and the individual credentials signature, but those could be different keys.
\item The issuer attaches the signed event to the metadata of a transaction and sends it to the blockchain.
\item The issuer gives to each user their corresponding signed credential along with their associated Merkle inclusion proof.
\end{itemize}

Likewise previous protocol events, all parties will process the transaction once confirmed, validate the event signature and timestamp the Merkle root with the time associated to the carrier transaction.

@@

In order to revoke a credentials, issuers have two alternatives:
\begin{enumerate}
\item They revoke all the credentials in a batch.
      This is done by signing a \emph{RevokeBatch} event with a \emph{revocation} key associated to the issuer DID.
      The signed event contains the Merkle root to revoke.
\item They revoke specific credentials associated to a batch.
      This is done by signing a \emph{RevokeCredentials} event with a \emph{revocation} key associated to the issuer DID. 
      The signed event contains both the Merkle root associated to the credentials to revoke, and the hashes of the specific credentials to revoke.
\end{enumerate}

@@

Once they receive credentials from issuers, subjects will present them to interested parties, called \emph{verifiers}. For example, a student may receive a verifiable credential from a university, and would like to present his credential to a potential employer. In our setting, the verifier will be a party following our protocol events from the blockchain. The steps to present and verify a credential are the following:
\begin{itemize}
\item We assume a safe communication channel between the subject and verifier.
      We also assume that the channel can be identified by an identifier $ch$.
\item The subject shares his credential and Merkle inclusion proof to the verifier.
\item The verifier then: 
	\begin{itemize}
	\item computes the Merkle root from the inclusion proof and the credential hash;
	\item extracts the issuer DID and signing key id from the credential claims,
	\item retrieves from the state he computed from the blockchain, the timestamps associated to the Merkle root and the issuing key, 
	\item checks if the Merkle root or credential hash has been revoked
	\item validates the issuer signature on the credential, and determines if it was signed at a time when the issuer key was valid.
	\end{itemize}
  \textbf{Note:} The signature of the IssueBatch event has already been verified by the protocol participants at the time of batch publication. 
\item At this point, the verifier knows that the credential was properly signed. Now, he shares a nonce to the subject and asks him to sign it with a key associated to his DID.
\item The subject signs the hash of $nonce || ch$ and returns the signature to the verifier.
\item The verifier now checks the signature and concludes that the credential subject is indeed the person presenting it.
\end{itemize}

@@

In a nutshell, the proposed protocol (see \figref{fig:e2econnector}) is composed
of two phases. In the first phase, the issuer shares a QR code with the holder.
This QR code contains the issuer's public key (or, maybe, its DID), a random
value ($code$, in \figref{fig:e2econnector}), and a session identifier ($token$,
in \figref{fig:e2econnector}). This exchange takes place through a secure (i.e.,
confidential and authenticated) channel, such as physical exchange. Probably
immediately, although also possibly at a later point in time, the prospective
holder sends back her own public key (or, alternatively, her DID) authenticated
with a MAC, where we use the random value
in the QR code as secret key. The issuer uses the session identifier to locate
the corresponding session and, now that it received in an authenticated manner
the public key of the holder, can leverage hybrid encryption techniques to
establish an end-to-end encrypted session, e.g., using ECIES \cite{abr01}.
Ensuring end-to-end encryption and authentication is important at this point, as
all the information exchanged between issuer and holder is routed through the
connector: an entity or set of entities who ensure that asynchronous
communication can take place between issuer and holder. Thus, ECIES is crucial
to ensure the secrecy of the information, end-to-end. Since the holder needs to
have certainty that the data it receives comes from the issuer (and ECIES does
not authenticate the sender), we have the issuer digitally sign the data it
sends through the ECIES-secured channel. Note that we could involve the random
value shared via the QR code for this. But that would force us to re-execute the
first phase of the protocol even when the QR code is compromised after having
completed the first phase of the protocol. Digital signatures using the issuer's
key pair prevent this.

@@

While the protocol is simple, and is based on common cryptographic constructions,
it is easy to get such protocols wrong, as pointed out in numerous occasions
\cite{davi01, kraw01}. Hence, we have used Tamarin%
\footnote{https://tamarin-prover.github.io/ (Last access, October 5th, 2021.)}
to model our protocol% \footnote{Source code for the model available here
  % \url{https://github.com/input-output-hk/atala-prism/blob/research-tamarin-connector/prism-backend/research/formalizations/connector.spthy}
  % \todo{Update to final URL.}}
, which is schematized in \figref{fig:e2econnector}
